%\subsection{\ac{PEATSA} Outputs}
%\label{sec:peatsa_outputs}

When a \ac{PEATSA} run is initiated, information is output to a number of places:

\begin{itemize}
	\item Standard output and standard error messages are written to the file specified at the command line in the \ac{PEATSA} execution call. (See Section~\ref{sec:executing_peatsa}.)
	\item Logging information is written to the file specified using the logfile option in the \ac{PEATSA} options file.
	\item A new directory is created in the location specified by the run\_name and working\_directory options in the \ac{PEATSA} options file.
\end{itemize}

\noindent The contents of the new directory created by \ac{PEATSA} are as followed:

\begin{itemize}
	\item docs/. This directory contains two types of files:
	\begin{itemize}
		\item PEATSAhistory.csv. Only a single file of this type is created, and that file is overwritten at the end of each \ac{PEATSA} iteration. This file contains extremely basic information about each \ac{PEATSA} iteration, such as how many \ac{EMTG} runs were executed, how many \ac{EMTG} runs found a feasible solution, and how many unique \ac{PEATSA} cases found a better value of the \ac{PEATSA} objective function in a given iteration. This information is useful for knowing if a \ac{PEATSA} run should be stopped, tweaked, and re-started (e.g., if no cases are producing feasible results) or if a \ac{PEATSA} run should be ended (e.g., because no cases have found more optimal solutions in several iterations).
		\item IterationX.csv, where X is the iteration index, starting with 0. A new file of this type is created at the end of each \ac{PEATSA} iteration. Each file gives information about the most optimal case that has been found up through that iteration for each unique \ac{PEATSA} case, as defined by the fingerprint. (See Section~\ref{sec:customizing_the_peatsa_options_file}.) Some information is given by default, such as the file path to the *.emtg and *.emtgopt file, and the feasibility index. Additional information is added by adding tuples to the extra\_csv\_column\_definitions list in the \ac{PEATSA} options file. (See Section~\ref{sec:customizing_the_peatsa_options_file}.)
	\end{itemize}
	\item results/. This directory contains subdirectories in which the output *.emtg, *.emtgopt, and *.emtg\_spacecraftopt files produced for each \ac{PEATSA} iteration are written.
	\item cases/. This directory contains subdirectories in which the input *.emtgopt files produced for each \ac{PEATSA} iteration are written.
	\item images/. This directory contains plots generated by the \ac{PEATSA} iterations.
	\item HardwareModels/. This directory contains \ac{EMTG} hardware files needed for each \ac{PEATSA} iteration (e.g., .emtg\_spacecraftopt).
	\item PEATSA\_iteration\_kill\_file.DEATH. The contents of the text file may be edited to end a \ac{PEATSA} iteration or an entire \ac{PEATSA} run. See the comments in the file for more details.
	\item PEATSA\_MidBake\_Options.py. The file is similar to a \ac{PEATSA} options file, but it specifically only has options that a user is allowed to change from iteration to iteration of a \ac{PEATSA} run.
	\item PEATSA\_ExecutedOptions.py. This file holds the values of the \ac{PEATSA} options used to kick off the \ac{PEATSA} run.
\end{itemize}